---
title: "KEGG/GO analysis"
output:
  html_notebook: default
  html_document: default
---

!!!change before running script:
1. name of organism in following chunk "convertKEGG_geneSymbol"

Here, we want to analyse the differentially expressed genes for differences in the KEGG and GO pathways

First we load all necessary libraries
```{r libraries}
.libPaths("/home/bioinf/bioinf_data/43_sovi/R/")

suppressPackageStartupMessages(library("KEGGREST"))
suppressPackageStartupMessages(library("GSA"))
suppressPackageStartupMessages(library(dplyr))
suppressPackageStartupMessages(library(tidyr))
suppressPackageStartupMessages(library(limma))
suppressPackageStartupMessages(library(edgeR))
suppressPackageStartupMessages(library(rtracklayer))
suppressPackageStartupMessages(library(reshape2))
suppressPackageStartupMessages(library("doParallel"))
suppressPackageStartupMessages(library(tximport))


```

Importing the .gmt files which were downloaded from the MsigDB homepage for the KEGG aswell as the GO analysis pathway. 
```{r gmtImport, echo=FALSE}

GO_gene_symbols <- GSA.read.gmt("../annotations/GO_All_c5.all.v6.0.symbols.gmt")
KEGG_gene_symbols <- GSA.read.gmt("../annotations/KEGG_c2.cp.kegg.v6.0.symbols.gmt")





```

Importing the RNAseq results
```{r importing_RNASeq}
tx2gene <- "/home/bioinf/bioinf_data/43_sovi/Projects/Metagenomics_pipeline/Model_data/e_coli/rnaseqworkflow-master/reference/Escherichia_coli_o157_h7_str_sakai.ASM886v1.cdna.ncrna_tx2gene.rds"
edgerres='/home/bioinf/bioinf_data/43_sovi/Projects/Metagenomics_pipeline/Model_data/e_coli/rnaseqworkflow-master/output/edgeR_dge.rds'
salmondir <- "/home/bioinf/bioinf_data/43_sovi/Projects/Metagenomics_pipeline/Model_data/e_coli/rnaseqworkflow-master/salmon"
metafile <- "/home/bioinf/bioinf_data/43_sovi/Projects/Metagenomics_pipeline/Model_data/e_coli/rnaseqworkflow-master/metadata.txt"

edgerres <- readRDS(edgerres)
tx2gene <- readRDS(tx2gene)

## Read metadata and reorder in the same order as the count matrix
metadata <- read.delim(metafile, header = TRUE, as.is = TRUE, sep = "\t")
(des <- model.matrix(~ time, data = metadata))


```

Convert KEGG_name into Genesysmbol
```{r convertKEGG_geneSymbol}

query_symbol_par <- rep("NA",times=length(edgerres$results$a2vsa5$symbol))

library("doParallel")
registerDoParallel(cores=15)

query_symbol_par <- foreach(i=1:length(edgerres$results$a2vsa5$symbol)) %dopar% {
    #print(i)
  tryCatch({
    query <- keggGet(paste("ecs:",edgerres$results$a2vsa5$symbol[i],sep = ""))
  },error=function(e){})
  
  tryCatch({
   query_symbol_par[i] <- query[[1]]$NAME
  },error=function(e){})
}

##create a vector with KEGGNames
query_symbol <- rep("NA",times=length(edgerres$results$a2vsa5$symbol))
for(i in 1:length(edgerres$results$a2vsa5$symbol)){
  
  if(!is.null(query_symbol_par[[i]])){query_symbol[i]<-query_symbol_par[[i]]}
  
}
names(query_symbol) <- edgerres$results$a2vsa5$symbol

```


Prepare KEGG analysis
```{r GSEA_KEGG_prep}

for(i in 1:length(KEGG_gene_symbols$geneset.names)){
  num_01<-num_01+1
  num<-0
  num_02 <-0
  for(k in 1:length(KEGG_gene_symbols$genesets[[i]])){
    if(sum(grepl(pattern=KEGG_gene_symbols$genesets[[i]][k],x=query_symbol,ignore.case = TRUE))!=0){
      num<-num+1
      KEGG_gene_symbols$genesets[[i]][k] <- grep(KEGG_gene_symbols$genesets[[i]][k],query_symbol,ignore.case = TRUE)
    }
    else{
      KEGG_gene_symbols$genesets[[i]][k] <- NA
    }
  }
}

omit_kegg_list <- lapply(KEGG_gene_symbols$genesets,function(x) x[!is.na(x)])
names(omit_kegg_list) <- KEGG_gene_symbols$geneset.names
omit_kegg_list <- omit_kegg_list[lapply(omit_kegg_list,length)>0]

```

##KEGG analysis
now we want to see which KEGG pathways are overrepresented
```{r GSEA_KEGG}
design <- cbind(Intercept=1,Group=c(0,1,0,1,0,1))
rownames(edgerres$data$counts) <- 1:length(rownames(edgerres$data$counts))

diff_KEGG <- camera(edgerres$data$counts,omit_kegg_list,design = design)
#View(diff_KEGG)

hist(diff_KEGG$PValue)
head(diff_KEGG)
```

##Scater plots
create scater plots for diagnostic purposes
```{r scater_prep}

##=============
##creating AnnotatedDataFrame which is a metadata/phenotype file
##=============
df <- data.frame(x=metadata$time,
                 row.names=metadata$ID)
metaData <-
  data.frame(labelDescription=c(
               "time"))

AnnotatedDataFrame()
AnnotatedDataFrame(data=df)
AnnotatedDataFrame(data=df, varMetadata=metaData)
as(df, "AnnotatedDataFrame")

obj <- AnnotatedDataFrame()
pData(obj) <- df
varMetadata(obj) <- metaData
validObject(obj)

##=============
##creating scater file
##=============

example_sceset <- newSCESet(countData = edgerres$data$counts, phenoData = obj)
example2 <- newSCESet(exprsData = log2(calculateCPM(example_sceset) + 1))

print("expression data")
exprs(example2)[1:3, 1:6]

print("count data")
exprs(example2)[1:3, 1:6]
```

##cummulative expression plot
This method plots the cumulative proportion of each cell’s library that is accounted for by the top highest-expressed features.
This plot gives an idea if the samples have entirely different expression distributions.
 
```{r cumulative_ expression_plot}

plot(example_sceset, exprs_values = "counts")
##add blocks if certain metadata can be added



dat <- (log10(calculateCPM(example_sceset) + 1))
dat2 <- melt(dat,id.vars="row")
ggplot(dat2, aes(x=Var2, y=value))+geom_boxplot()+
guides(fill=FALSE)+
  scale_y_continuous(name="log10(Cpm)+1")+
  labs(title="boxplot of expressions of different samples")

```

##Exression plots
```{r ExprPlots}
plotExpression(example_sceset, rownames(example_sceset)[1:6],x = "x", exprs_values = "exprs")

```

##Quality control
```{r QCplots}

example_sceset <- calculateQCMetrics(example_sceset)
varLabels(example_sceset)

keep_feature <- rowSums(counts(example_sceset) > 0) > 4
example_sceset <- example_sceset[keep_feature,]
## Plot QC
plotQC(example_sceset, type = "highest-expression", exprs_values = "counts")

plotQC(example_sceset, type = "exprs-freq-vs-mean")

plotFeatureData(example_sceset, aes(x = n_cells_exprs, y = pct_total_counts))

example_sceset_pca <- plotPCA(example_sceset, pca_data_input = "pdata", 
                          detect_outliers = TRUE, return_SCESet = TRUE)

plotPCA(example_sceset, exprs_values = "counts",colour_by = "x")
```

